---
title: "LogCheck"
author: "Chuan Hong"
date: "3/16/2017"
output:
html_document: default
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

## Data import and munge

```{r}
library(car)

POST <- read.csv("REVISED_FINAL_POST_OM - PRE-QREMOVAL.csv",header = TRUE,sep = ",",stringsAsFactors = FALSE)
PRE <- read.csv("REVISED_FINAL_PRE_OM - PRE-QREMOVAL.csv",header = TRUE,sep = ",",stringsAsFactors = FALSE)

Post_proc <- lapply(POST, FUN = function(foo) car::recode(foo, "'Never' = 0 ; 'Annually' = 1 ; 'Seasonally' = 2 ; 'Monthly' = 3 ; 'Weekly' = 4 ; 'Daily' = 5 ; 'Not applicable' = NA ; 'Not Applicable' = NA ; 'No logging' = NA ; 'No understanding' = NA ; '' = NA ; 'Y' = T", as.numeric.result = T, as.factor.result = F))
Post_proc <- data.frame(Post_proc)
IsExperimental = Post_proc[2:nrow(Post_proc),c(1,ncol(Post_proc))]
cutdown <- ncol(Post_proc)-3
Post_proc <- Post_proc[,1:cutdown]
best_prac <- as.vector(Post_proc[1,2:cutdown],mode='numeric')
best_prac = append(1, best_prac)
Post_proc <- Post_proc[2:nrow(Post_proc), ]

Pre_proc <- lapply(PRE, FUN = function(foo) car::recode(foo, "'Never' = 0 ; 'never' = 0 ; 'Annually' = 1 ; 'Seasonally' = 2 ; 'Monthly' = 3 ; 'Weekly' = 4 ; 'Daily' = 5 ; 'Not applicable' = NA ; 'Not Applicable' = NA ; 'Multiple answer' = NA ; 'Mulitple answer' = NA ; 'As needed' = NA ; 'No logging' = NA ; 'No understanding' = NA ; 'NR' = NA ; '' = NA", as.numeric.result = T, as.factor.result = F))
Pre_proc <- data.frame(Pre_proc)
cutdown <- ncol(Pre_proc)-2
Pre_proc <- Pre_proc[2:nrow(Pre_proc),1:cutdown]

Post_proc2 <- subset(Post_proc, BPL.BLD.ID %in% Pre_proc$BPL.BLD.ID)
Pre_proc2 <- subset(Pre_proc, BPL.BLD.ID %in% Post_proc$BPL.BLD.ID)
Post_score <- data.frame(mapply(`/`,Post_proc2,best_prac))
Pre_score <- data.frame(mapply(`/`,Pre_proc2,best_prac))
Pre_score <- Pre_score[with(Pre_score, order(BPL.BLD.ID)), ]
Post_score <- Post_score[with(Post_score, order(BPL.BLD.ID)), ]
Pre_score[1000 > Pre_score & Pre_score > 1] <- 1
Post_score[1000 > Post_score & Post_score > 1] <- 1

Pre_post = Post_score - Pre_score
Pre_post[,1] = Pre_score[,1]

IsExperimental <- lapply(POST[2:nrow(POST),c(1,ncol(POST))], FUN = function(foo) car::recode(foo, "'' = F ; 'Y' = T", as.numeric.result = T, as.factor.result = F))
IsExperimental <- data.frame(IsExperimental)
IsExperimental <- subset(IsExperimental, BPL.BLD.ID %in% Pre_score$BPL.BLD.ID)
IsExperimental <- IsExperimental[with(IsExperimental, order(BPL.BLD.ID)), ]
```


## Load libraries and data

```{r}

library(data.table)
library(dplyr)
library(ggplot2)

logcheck.usage <- fread(input = "./LOGCHECK USERS WITH POST.csv", 
                        header = TRUE)

# histogram
hist(logcheck.usage$LogCheck.Usage, ylab ="Frequency", xlab
="Usage", main = "Average LogCheck Usage")

stem(logcheck.usage$LogCheck.Usage)

```

#### - Merge Pre_post and IsExperimental, and list TRUE and FALSE
```{r}
# Merge Pre_post and IsExperimental
pre_post.isexp <- merge(Pre_post,IsExperimental,by='BPL.BLD.ID', all.x = TRUE)

summary(pre_post.isexp$LogCheck)

```

#Do filtering

```{r}

deleted_behaviors <- c()
Pre_score.filtered <- Pre_score
for(i in colnames(Pre_score[,2:ncol(Pre_score)])){
isone <- na.omit(Pre_score[[i]]) == 1
a <- ftable(isone)
if (a[2]/sum(a) > 0.75 | all(isone)){
  deleted_behaviors <- c(deleted_behaviors,i)
  pre_post.isexp[, i] <- NULL
  Pre_score.filtered[, i] <- NULL}}

deleted_behaviors
```
### - Do histograms of pre scores
```{r pre hist}
prehist <- vector(mode="list", length=5)
names(prehist) <- c("histogram", "maxfreq", "ndigits", "multiplier", "rounded")
par(mfrow=c(1,2))
prehist$histogram <- hist(unlist(Pre_score[,2:ncol(Pre_score)]), plot=FALSE)
prehist$maxfreq <- max(prehist$histogram$counts)
prehist$ndigits <- nchar(as.character(prehist$maxfreq))
prehist$multiplier <- 10^(prehist$ndigits-2)
prehist$rounded <- ceiling(prehist$maxfreq/prehist$multiplier)*prehist$multiplier
plot(prehist$histogram, main="(A) Scores Before Filtering", xlab="Behavior Compliance Score",ylim = c(0, prehist$rounded))
prehist2 <- hist(unlist(Pre_score.filtered[,2:ncol(Pre_score.filtered)]), main="(B) Scores After Filtering", xlab="Behavior Compliance Score", ylim = c(0, prehist$rounded))
```
### - Create new df call pre_post.all with all info: usage, isExperimental, group mean (X1, X2, ..., X8), grand mean (Xall)
```{r}
pre_post.all <- pre_post.isexp %>%
  mutate(X2 = rowMeans(select(pre_post.isexp, starts_with("X2")), na.rm = TRUE),
         X3 = rowMeans(select(pre_post.isexp, starts_with("X3")), na.rm = TRUE),
         X4 = rowMeans(select(pre_post.isexp, starts_with("X4")), na.rm = TRUE),
         X5 = rowMeans(select(pre_post.isexp, starts_with("X5")), na.rm = TRUE),
         X6 = rowMeans(select(pre_post.isexp, starts_with("X6")), na.rm = TRUE),
         X7 = rowMeans(select(pre_post.isexp, starts_with("X7")), na.rm = TRUE),
         X8 = rowMeans(select(pre_post.isexp, starts_with("X8")), na.rm = TRUE),
         Xall = rowMeans(select(pre_post.isexp, starts_with("X")), na.rm = TRUE))
```

## T-Tests

```{r t_test}
ttest <- function(coln,group){
  col_C <- coln[group == 'FALSE']
  col_E <- coln[group == 'TRUE']
  t.test(col_E,col_C,alternative = 'greater')$p.value
}

ttest.p <- sapply(pre_post.all %>% select(-BPL.BLD.ID, -LogCheck), 
                  function(x) ttest(x, pre_post.isexp$LogCheck))

hist(ttest.p,width=5)
signif_p <- Filter(function(x) x < .05, ttest.p)
signif_p
```

## Wilcoxon-Mann-Whitney tests

```{r wmw_test}
#create function for WMW
wmwtest <- function(coln, group) {
  col_C <- coln[group == 'FALSE']
  col_E <- coln[group == 'TRUE']
  wilcox.test(col_E, col_C, alternative = 'greater')$p.value
}

#conduct wmw test and save p values as a dataframe 
wmw.p <- sapply(pre_post.all %>% select(-BPL.BLD.ID, -LogCheck), 
                wmwtest, group=pre_post.all$LogCheck)

ggplot(data.frame(name = names(wmw.p), p_val = wmw.p),aes(p_val))+
  geom_histogram()+
  theme_bw()+
  ggtitle("Histogram of Wilcoxon-Mann-Whitney test p value")

```
## Adjust P-values and put in dataframe
```{r fdr}
wmw.q <- p.adjust(wmw.p, method = "BH")
ttest.q <- p.adjust(ttest.p, method = "BH")

library(fdrtool)
wmw.q2 <- fdrtool(wmw.p[complete.cases(wmw.p)], statistic = "pvalue")
ttest.q2 <- fdrtool(ttest.p[complete.cases(ttest.p)], statistic = "pvalue")

pvals <- merge(data.frame(name = names(ttest.p), tp_val = ttest.p, tq_val = ttest.q, tfdr2 = ttest.q),data.frame(name = names(wmw.p), wp_val = wmw.p, wq_val = wmw.q, wfdr2 = wmw.q),by='name', all.x = TRUE)
pvals.sorted <- pvals[with(pvals, order(tp_val, wp_val)), ]
pvals.sorted[pvals.sorted$tp_val <= 0.1,]

```
## make p value table
```{r}
a <- pvals[pvals$tp_val <= 0.05,c(1,2,3,5,6)]
a[,2:5] <- round(a[,2:5],digits=4)
a <- t(a)
```
### 3) Table of all means, SDs and SEMs
### 4) Table of all mean differences

```{r}
library(tibble)
library(dplyr)

sem <- function(x, ...) {
  sd(x, ...)/sqrt(length(na.omit(x)))
}

pre_post.stat <- pre_post.all %>%
  group_by(LogCheck) %>%
  summarise_at(vars(matches("X")), funs(mean, sd, sem), na.rm = TRUE) %>%
  mutate(logc = ifelse(LogCheck == TRUE, "LogCheck", "Control")) %>%
  column_to_rownames(var="logc") %>%
  select(-LogCheck)
```

```{r}
library(data.table)
library(tidyr)

pre_post.stat.rshaped <- as.data.frame(t(pre_post.stat)) %>% 
  rownames_to_column("Behavior")  %>%
  separate(col = Behavior, into = c("Behavior", "measure"), sep = "_") %>%
  setDT() %>%
  dcast(Behavior ~ measure, value.var = c("LogCheck", "Control")) %>%
  mutate(mean_diff = LogCheck_mean - Control_mean)

pre_post.stat.rshaped #Show all means, SDs, SEMs, mean differences in Exp and Ctl group
orderedbehaviors <- pre_post.stat.rshaped[with(pre_post.stat.rshaped, order(-mean_diff)), c(1,8)]

library(ggplot2)

pre_post.stat.mean <- pre_post.stat.rshaped %>% 
  select(Behavior, LogCheck_mean, Control_mean, LogCheck_sem, Control_sem) %>%
  gather(LogCheck_mean, Control_mean, key = "group", value = "mean") %>%
  mutate(sem = ifelse(group == "LogCheck_mean", LogCheck_sem, Control_sem),
         group = sub("([A-Z][a-z]+)_.+", "\\1", group)) %>%
  select(-LogCheck_sem, -Control_sem) %>%
  merge(data.frame(Behavior = names(ttest.p), p_val = ttest.p),by='Behavior', all.x = TRUE) %>%
  mutate(sig_level = ifelse(p_val > .1, "", ifelse(p_val > .05, "*", "**")))

```

## Make df for beanplotting

```{r}

foruse <- pre_post.all %>% 
  mutate(Group = ifelse(LogCheck == TRUE, "LogCheck", "Control")) %>%
  subset(select = -LogCheck)
plotframe <- data.frame(BPL.BLD.ID=integer(),
                 Improvement=double(), 
                 Behavior=character(), 
                 Group=character(),
                 stringsAsFactors=FALSE)
for (behavior in orderedbehaviors$Behavior){
  newdf <- filter_(foruse,paste(behavior," > -10"))
  newdf <- newdf[,c("BPL.BLD.ID", behavior, "Group")]
  names(newdf)[names(newdf)==behavior] <- "Improvement"
  newdf$Behavior <- behavior
  plotframe <- rbind(plotframe, newdf)}


lvl <- orderedbehaviors$Behavior
lvl <- lvl[length(lvl):1]
plotframe$Behavior <- factor(plotframe$Behavior, levels=lvl)

```

#Do beanplot
```{r}
library(beanplot)
library(grid)
library(gplots)

lvl <- orderedbehaviors$Behavior
lvl <- lvl[length(lvl):1]
pre_post.stat.mean.ordered <- pre_post.stat.mean
pre_post.stat.mean.ordered <- within(pre_post.stat.mean.ordered, Behavior <- factor(Behavior, levels=lvl))

plot1 <- pre_post.stat.mean.ordered %>%
  ggplot(aes(x = Behavior, y = mean, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values=col2hex(c("purple","orange"))) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), position = position_dodge(0.9)) +
  geom_text(aes(label=sig_level), position = position_dodge(0.9)) +
  theme_bw()+coord_flip()+theme(legend.position="bottom",axis.title.x=element_blank(),
                                axis.title.y=element_blank())
  #ggtitle("Paired bar chart for all individual Behaviors with SEM whiskers")

vp.BottomRight <- viewport(height=unit(0.875, "npc"),
                           width=unit(0.5, "npc"), 
                           just=c("left","top"), 
                           y=0.925, x=0.45)

par(mfrow=c(1,2))

beanplot(Improvement ~ reorder(Group, Improvement, mean) * Behavior, bw="nrd0", data = plotframe,
         side = "b", horizontal = TRUE, yaxt = "n", las=1, col = list("purple", "orange"), border = c("purple2","darkorange"), ylim = c(-1.5,1.5), yaxs = "i")
axis(4, at=1:length(lvl), labels=FALSE)

#ylab = "Behavior", 
#main = "Improvement on Each Behavior", xlab = "Improvement", 
#legend("bottomright", bty="n",c("Control", "LogCheck"), fill = c("purple", "orange"))

print(plot1, vp=vp.BottomRight)

```

### 2) Paired bar chart with SEM whiskers or significant and marginally significant results -- a separate one for groups, and markings for significance and high significance. (Bar chart should have means.)

```{r}

pre_post.stat.mean %>%
  filter(grepl("^X\\d$", Behavior)) %>%
  ggplot(aes(x = Behavior, y = mean, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values=col2hex(c("purple","orange"))) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), 
                position = position_dodge(0.9),
                width =.5) +
  geom_text(aes(label=sig_level),position = position_dodge(0.9))+
  theme_bw()+theme(legend.position="bottom") +
  ylab("Mean Improvement")+
  ggtitle("(B) Improvement on Aggregate Behaviors")

# Barchart for individuals and groups with p < 0.05

pre_post.stat.mean %>%
  filter(Behavior %in% as.character(pvals[pvals$tp_val < 0.05, ]$name)) %>%
  ggplot(aes(x = Behavior, y = mean, fill = group)) +
  scale_fill_manual(values=col2hex(c("purple","orange"))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), 
                position = position_dodge(0.9),
                width = .5) +
  theme_bw()+theme(legend.position="bottom")+
  ylab("Mean Improvement")+
  ggtitle("Behaviors with Significant (p < 0.05) Improvement")

# Barchart for all

pre_post.stat.mean %>%
  filter(grepl("^Xall$", Behavior)) %>%
  ggplot(aes(x = Behavior, y = mean, fill = group)) +
  scale_fill_manual(values=col2hex(c("purple","orange"))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), 
                position = position_dodge(0.9),
                width = .5) +
  theme_bw()+#theme(legend.position="bottom")+
  ylab("Mean Improvement")+
  ggtitle("(A) Overall Improvement on Tracked Behaviors")
```

```{r, sumstats}

library(psych)

pre_post.sumstats <- describe(pre_post)
write.csv(pre_post.sumstats, "pre_post.sumstats.csv")

````
